---
title: "Descriptive and univariate analyses"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, echo = FALSE, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tidyverse)
library(readr)
library(dplyr)
library(ggpubr)
library(patchwork)
library(sjmisc)
library(sjlabelled)
library(tableone)
library(stringr)
library(kableExtra)
library(nlme)
library(lme4)
library(gamlj)

# NOTES
# bl_int_4 = Education levels (re-coded for ease)
# 1 = no education, 2 = primary, 3 = secondary, 4 = university, 5 = grad school, 6 = other

# bl_int_2 = Marital status
## 1 = Single; 2 = Married/living with partner; 3 = Separated or divorced; 4 = Widow or widower

# bl_int_6 = Ethnicity
# Aboriginal - 1, Arab - 2, African - 3, Black - 4, Chinese - 5, Filipino - 6, Hispanic - 7, Jamaican - 8, Japanese - 9, Korean - 10, Metis - 11, White - 12, West Asian (Iranian, Afghan, etc.) - 13, Southeast Asian (Vietnamese, Cambodian, etc.) - 14, South Asian (Sri Lankan, East Indian, etc.) - 15, Other: - 16, Refused - 17, Don`t know - 18

# 1 = White/Caucasian = 12
# 2 = Black, Caribbean, African = 3, 4, 8 
# 3 = Asian = 5, 6, 9, 10, 13, 14, 15
# 4 = Hispanic = 7
# 5 = Native = 1, 11
# 6 = Other/multi = 2, 16, 17, 18 or >18

# bl_int_9 = Employment
# 1 = Working full time; 2 = Working part time; 3 = Temporarily laid off; 4 = Unemployed
# 5 = Retired; 6 = Homemaker; 7 = Student; 8 = On disability; 9 = Volunteering 
# 10 = Working with modified tasks; 11, On leave with pay; 12 = On leave without pay; 13 = On welfare

# NEED clin_path_8 for tumor grade

```

```{r echo=FALSE, message=FALSE}
data <- read_csv("data/mtl_scored_apr2023.csv") %>%
  filter(gender %in% c("FEMALE", "MALE")) %>%
  filter(age_y > 15) %>%
  mutate(bl_int_6 = str_replace(bl_int_6, "\\;", ""),
         bl_int_4 = factor(case_when(bl_int_4 == 1 ~ 1,
                                     bl_int_4 == 2 ~ 2,
                                     bl_int_4 == 3 | bl_int_4 == 4 | bl_int_4 == 5 | bl_int_4 == 6 | bl_int_4 == 7 | bl_int_4 == 8 ~ 3,
                                     bl_int_4 == 9 | bl_int_4 == 10 | bl_int_4 == 11 ~ 4,
                                     bl_int_4 == 12 | bl_int_4 == 13 ~ 5,
                                     bl_int_4 == 14 ~ 6)),
         education = case_when(bl_int_4 == "1" ~ "No education",
                               bl_int_4 == "2" ~ "Primary",
                               bl_int_4 == "3" ~ "Secondary",
                               bl_int_4 == "4" ~ "University",
                               bl_int_4 == "5" ~ "Grad school",
                               bl_int_4 == "6" ~ "Other"),
         education_num = as.numeric(bl_int_4),
         bl_int_6 = factor(case_when(bl_int_6 == 12 ~ 1,
                                     bl_int_6 == 3 | bl_int_6 == 4 | bl_int_6 == 8 ~ 2,
                                     bl_int_6 == 5 | bl_int_6 == 6 | bl_int_6 == 9 |
                                       bl_int_6 == 10 | bl_int_6 == 13 |bl_int_6 == 14 |
                                       bl_int_6 == 15 ~ 3,
                                     bl_int_6 == 7 ~ 4,
                                     bl_int_6 == 1 | bl_int_6 == 11 ~ 5,
                                     bl_int_6 == 2 | bl_int_6 > 15 ~ 6)),
         ethnicity = case_when(bl_int_6 == "1" ~ "Caucasian/White",
                               bl_int_6 == "2" ~ "Black/Carribbean/African",
                               bl_int_6 == "3" ~ "Asian",
                               bl_int_6 == "4" ~ "Hispanic",
                               bl_int_6 == "5" ~ "Native",
                               bl_int_6 == "6" ~ "Other/Multiple"),
         marital = case_when(bl_int_2 == 1 ~ "Single",
                             bl_int_2 == 2 ~ "Married/Cohabiting",
                             bl_int_2 == 3 ~ "Separated/Divorced",
                             bl_int_2 == 4 ~ "Widowed"),
         bl_deceased = case_when(bl_deceased == "No" ~ "No",
                                 bl_deceased == "Unknown" ~ "No",
                                 bl_deceased == "Yes" ~ "Yes"),
         tumor_location = case_when(bl_le_tessscore_sc >= 0 ~ "lower",
                              bl_ue_tessscore_sc >= 0 ~ "upper"),
         tumor_grade = case_when(clin_path_8 == "0" ~ 0,
                                 clin_path_8 == "1" ~ 1,
                                 clin_path_8 == "2" ~ 2,
                                 clin_path_8 == "3" ~ 3,
                                 clin_path_8 == "4" ~ 4,
                                 clin_path_8 == "Unknown" ~ as.numeric(NA))) %>%
  mutate(change_function = ifelse(bl_agreement_function == m12_agreement_function, "No change", "Change"),
         change_pain = ifelse(bl_agreement_pain == m12_agreement_pain, "No change", "Change")) %>%
  mutate(change_function_direction = case_when(bl_level_agreement_function_dummy == m12_level_agreement_function_dummy ~ "No change",
                                               (bl_level_agreement_function_dummy == "MD higher" | bl_level_agreement_function_dummy == "MD lower") & m12_level_agreement_function_dummy == "neutral" ~ "Improved",
                                               bl_level_agreement_function_dummy == "neutral" & (m12_level_agreement_function_dummy == "MD lower" | m12_level_agreement_function_dummy == "MD higher") ~ "Worsened"),
         change_pain_direction = case_when(bl_level_agreement_pain_dummy == m12_level_agreement_pain_dummy ~ "No change",
                                               (bl_level_agreement_pain_dummy == "MD higher" | bl_level_agreement_pain_dummy == "MD lower") & m12_level_agreement_pain_dummy == "neutral" ~ "Improved",
                                               bl_level_agreement_pain_dummy == "neutral" & (m12_level_agreement_pain_dummy == "MD lower" | m12_level_agreement_pain_dummy == "MD higher") ~ "Worsened"))

```

## Notes

-   Only cases diagnosed at age 16 or later are included (consent)
-   clin_path_8 = tumor grade
-   tumor type = 1 Soft tissue, 2 Bone
-   bl_deceased = passed before surgical intervention

## SAMPLE
```{r echo=FALSE, warning=FALSE, message=FALSE}
data_sample <- data %>%
  filter(!is.na(bl_assess_complete))

var_sample <- c("gender", "age_y", "bl_tumour_type", "tumor_location", "tumor_grade", "marital", "education", "ethnicity", "bl_deceased", "bl_optimism", "bl_pessimism", "bl_tess", "bl_facit_score", "bl_qol_7")

# Determine which variables are categorical
cat_vars_sample <- c("gender", "bl_tumour_type", "tumor_location", "tumor_grade","marital", "education", "ethnicity", "bl_deceased")

#All cases table
table_sample <- CreateTableOne(vars = var_sample, 
                          factorVars = cat_vars_sample,
                          data = data_sample)
table_sample
```

### Analyses

#### Table for differences in agreement on function based on demographic and medical factors

AGREEMENT on Function

Stratification by: 1 = Patient and doctor agree 0 = Patient and doctor
disagree

Agreement on function was associated with tumor grade, whether or not
patient passed before surgery (also related to tumor grade = more
invasive cancer), with better baseline function, fatigue, QOL.

```{r echo=FALSE}
univariate <- data %>%
  dplyr::select(subject_id, gender, 
                procedure, tumor_location, tumor_grade,
                area, bl_prez_5_1, bl_tess, bl_msts_extremity,
                bl_deceased, age_y, bl_tumour_type,
                bl_optimism, bl_pessimism,
                bl_int_9, bl_int_11,
                marital, education, ethnicity,
                bl_tess, bl_facit_score, bl_qol_7,
                bl_ioe_recover_time, bl_ioe_complication, bl_ioe_recover_difficulty,
                bl_agreement_function, bl_agreement_pain,
                m12_agreement_function, m12_agreement_pain,
                change_function, change_pain) 


# Create a list of variables to be looked at
var_list_1 <- c("gender", "age_y", "bl_tumour_type", "tumor_location", "tumor_grade", "marital", "education", "ethnicity", "bl_deceased", "bl_optimism", "bl_pessimism", "bl_tess", "bl_facit_score", "bl_qol_7")

# Determine which variables are categorical
cat_vars_1 <- c("gender", "bl_tumour_type", "tumor_location", "tumor_grade","marital", "education", "ethnicity", "bl_deceased")

#All cases table
table_1 <- CreateTableOne(vars = var_list_1, 
                          factorVars = cat_vars_1,
                          data = univariate,
                          strata = c("bl_agreement_function"))
table_1
```

#### Table for differences in agreement on pain based on demographic and medical factors

AGREEMENT on Pain

Stratification by: 1 = Patient and doctor agree 0 = Patient and doctor
disagree

Agreement with doctors on pain at baseline was associated with tumor
type, patient's optimism, higher baseline level of self-reported
functioning, quality of life and lower fatigue.

```{r echo=FALSE}
# Create a list of variables to be looked at
var_list_2 <- c("gender", "age_y", "bl_tumour_type", "tumor_location", "tumor_grade","marital", "education", "ethnicity", "bl_deceased", "bl_optimism", "bl_pessimism", "bl_tess", "bl_facit_score", "bl_qol_7")

# Determine which variables are categorical
cat_vars_2 <- c("gender", "bl_tumour_type", "tumor_location", "tumor_grade", "marital", "education", "ethnicity", "bl_deceased")

#All cases table
table_2 <- CreateTableOne(vars = var_list_2, 
                          factorVars = cat_vars_2,
                          data = univariate,
                          strata = c("bl_agreement_pain"))
table_2
```

### Change in function from baseline to 1 year

There is a very small correlation between baseline and 1 year agreement
on function. Dummy variable suggests that 58% see no change, but 42%
demonstrated change in agreement on function.

```{r echo=FALSE, warning=FALSE, message=FALSE}
data %>%
  dplyr::select(change_function) %>%
  group_by(change_function) %>%
  count() %>%
  drop_na() %>%
  ungroup() %>%
  mutate(sum = sum(n),
         Prop = (prop = round(n/sum,2)*100)) %>%
  dplyr::select(-sum) %>%
  kable() %>%
  kable_styling(full_width = FALSE)

ggscatter(data, x = "bl_level_agreement_function", y = "m12_level_agreement_function",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   ) + 
  stat_cor(method = "spearman", 
           label.x = 2, label.y = 4,
           p.accuracy = 0.001, r.accuracy = 0.01) +
  geom_jitter()
```

## Variables associated with change on function over time

No demographic or medical variables was associated with change in
agreement on function.

```{r echo=FALSE}
# var_list_3 <- c("gender", "age_y", "bl_tumour_type", "tumor_location", "clin_path_8", "marital", "education", "ethnicity", "bl_deceased", "bl_optimism", "bl_pessimism", "bl_tess", "bl_facit_score", "bl_qol_7")
# 
# # Determine which variables are categorical
# cat_vars_3 <- c("gender", "bl_tumour_type", "tumor_location", "clin_path_8","marital", "education", "ethnicity", "bl_deceased")
# 
# #All cases table
# table_3 <- CreateTableOne(vars = var_list_3, 
#                           factorVars = cat_vars_3,
#                           data = univariate,
#                           strata = c("change_function"))
# table_3
```

### Change in pain from baseline to 1 year

There is literally no correlation between baseline and 1 year agreement
suggesting that some people agree more and others disagree more. Dummy
variably suggests that 55% see no change, but 45% demonstrated change in
agreement on pain.

```{r echo=FALSE, warning=FALSE, message=FALSE}
data %>%
  dplyr::select(change_pain) %>%
  group_by(change_pain) %>%
  count() %>%
  drop_na() %>%
  ungroup() %>%
  mutate(sum = sum(n),
         Prop = (prop = round(n/sum,2)*100)) %>%
  dplyr::select(-sum) %>%
  kable() %>%
  kable_styling(full_width = FALSE)

ggscatter(data, x = "bl_level_agreement_pain", y = "m12_level_agreement_pain",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   ) + 
  stat_cor(method = "spearman", 
           label.x = 2, label.y = 4,
           p.accuracy = 0.001, r.accuracy = 0.01) +
  geom_jitter()
```

## Variables associated with change on pain over time

No demographic variables were associated with change in agreement on
pain.

```{r echo= FALSE}
# var_list_4 <- c("gender", "age_y", "bl_tumour_type", "tumor_location", "clin_path_8", "marital", "education", "ethnicity", "bl_deceased", "bl_optimism", "bl_pessimism", "bl_tess", "bl_facit_score", "bl_qol_7")
# 
# # Determine which variables are categorical
# cat_vars_4 <- c("gender", "bl_tumour_type", "tumor_location", "clin_path_8","marital", "education", "ethnicity", "bl_deceased")
# 
# #All cases table
# table_4 <- CreateTableOne(vars = var_list_4, 
#                           factorVars = cat_vars_4,
#                           data = univariate,
#                           strata = c("change_pain"))
# table_4
```

## Outcomes over time based on agreement on function

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=11}
data_over_time <- data %>%
  dplyr::select(subject_id, age_y, bl_level_agreement_pain, bl_level_agreement_function, bl_agreement_function, bl_agreement_pain, bl_level_agreement_pain_dummy, bl_level_agreement_function_dummy, bl_tess, m12_tess, bl_qol_7, m12_qol_7, bl_facit_score, m12_facit_score) %>%
  drop_na(bl_agreement_function) %>%
  pivot_longer(cols = c(bl_tess, m12_tess, bl_qol_7, m12_qol_7, bl_facit_score, m12_facit_score),
               names_to = c("time", "measure"),
               names_sep = "_",
               values_to = "score")
# data_over_time %>% 
#   ggplot(aes(x = time, y = score)) +
#   geom_boxplot(aes(fill = factor(bl_level_agreement_function))) +
#   facet_grid(~measure)

data_over_time %>%
  ggplot(aes(x = time, y = score)) +
  geom_boxplot(aes(fill = factor(bl_level_agreement_function_dummy))) +
  facet_grid(~measure)
```

## Outcomes over time based on agreement on pain

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=11}
data_over_time <- data %>%
  dplyr::select(subject_id, age_y, bl_level_agreement_pain, bl_level_agreement_function, bl_agreement_function, bl_agreement_pain, bl_level_agreement_function_dummy, bl_level_agreement_pain_dummy, bl_tess, m12_tess, bl_qol_7, m12_qol_7, bl_facit_score, m12_facit_score) %>%
  drop_na(bl_agreement_function) %>%
  pivot_longer(cols = c(bl_tess, m12_tess, bl_qol_7, m12_qol_7, bl_facit_score, m12_facit_score),
               names_to = c("time", "measure"),
               names_sep = "_",
               values_to = "score")
# data_over_time %>% 
#   drop_na(bl_level_agreement_pain) %>%
#   ggplot(aes(x = time, y = score)) +
#   geom_boxplot(aes(fill = factor(bl_level_agreement_pain))) +
#   facet_grid(~measure)

data_over_time %>% 
  drop_na(bl_level_agreement_pain_dummy) %>%
  ggplot(aes(x = time, y = score)) +
  geom_boxplot(aes(fill = factor(bl_level_agreement_pain_dummy))) +
  facet_grid(~measure)
```

## Change in agreement on function and relationship to main outcomes

```{r echo=FALSE, warning=FALSE, message=FALSE}
var_list_change <- c("m12_tess", "m12_qol_7", "m12_facit_score")

#All cases table
table_change_function <- CreateTableOne(vars = var_list_change, 
                               data = data,
                               strata = c("change_function"))
table_change_function
```

## Change in agreement on pain and relationship to main outcomes

```{r echo=FALSE, warning=FALSE, message=FALSE}
#All cases table
table_change_pain <- CreateTableOne(vars = var_list_change, 
                               data = data,
                               strata = c("change_pain"))
table_change_pain
```

# Demographic factors and correlations to the main outcomes

Self-reported function at 12 months was associated with age, education,
being widowed, tumor grade.

Self-reported quality of life at 12 months was associated with education
level.

Fatigue at 12 months was associated with education and tumor grade.

```{r echo=FALSE, message=FALSE, results=FALSE}
m_tess_gender <- lm(m12_tess ~ gender, data = data)
summary(m_tess_gender) #no

m_tess_age <- lm(m12_tess ~ age_y, data = data)
summary(m_tess_age) #yes

m_tess_marital <- lm(m12_tess ~ marital, data = data)
summary(m_tess_marital) #widowed

m_tess_education <- lm(m12_tess ~ education_num, data = data)
summary(m_tess_education) # yes lower education

m_tess_ethnicity <- lm(m12_tess ~ ethnicity, data = data)
summary(m_tess_ethnicity) #NO

m_tess_tumor <- lm(m12_tess ~ bl_tumour_type, data = data)
summary(m_tess_tumor) #yes for the 2 types (n=3)

m_tess_grade <- lm(m12_tess ~ tumor_grade, data = data)
summary(m_tess_grade) #yes

###
m_qol_gender <- lm(m12_qol_7 ~ gender, data = data)
summary(m_qol_gender) #no

m_qol_age <- lm(m12_qol_7 ~ age_y, data = data)
summary(m_qol_age) #no

m_qol_marital <- lm(m12_qol_7 ~ marital, data = data)
summary(m_qol_marital) #no

m_qol_education <- lm(m12_qol_7 ~ education, data = data)
summary(m_qol_education) # yes lower education, minimum

m_qol_ethnicity <- lm(m12_qol_7 ~ ethnicity, data = data)
summary(m_qol_ethnicity) #NO

m_qol_tumor <- lm(m12_qol_7 ~ bl_tumour_type, data = data)
summary(m_qol_tumor) #no

m_qol_grade <- lm(m12_qol_7 ~ tumor_grade, data = data)
summary(m_qol_grade) #yes

###
m_facit_gender <- lm(m12_facit_score ~ gender, data = data)
summary(m_facit_gender) #no

m_facit_age <- lm(m12_facit_score ~ age_y, data = data)
summary(m_facit_age) #no

m_facit_marital <- lm(m12_facit_score ~ marital, data = data)
summary(m_facit_marital) #no

m_facit_education <- lm(m12_facit_score ~ education, data = data)
summary(m_facit_education) # yes lower education, minimum

m_facit_ethnicity <- lm(m12_facit_score ~ ethnicity, data = data)
summary(m_facit_ethnicity) #NO

m_facit_tumor <- lm(m12_facit_score ~ bl_tumour_type, data = data)
summary(m_facit_tumor) #no

m_facit_grade <- lm(m12_facit_score ~ tumor_grade, data = data)
summary(m_facit_grade) #yes
```

## UNIVARIATE: OUTCOME = TESS (Self-reported function)

Notes: cases that passed before the intervention removed (N \~ 107)

### based on agreement on function

```{r echo=FALSE, fig.width=8}
data_mixed_change_tess <- data %>%
  filter(bl_deceased != "Yes") %>%
  dplyr::select(subject_id, gender, age_y, education_num, marital, tumor_grade, change_function_direction, change_pain_direction, bl_level_agreement_pain, bl_level_agreement_function, bl_agreement_function, bl_agreement_pain, bl_level_agreement_function_dummy, bl_level_agreement_pain_dummy, 
                bl_tess, m12_tess) %>%
  pivot_longer(cols = c(bl_tess, m12_tess),
               names_to = "time",
               values_to = "score") %>%
  mutate(subject_id = as.numeric(subject_id),
         time = as.factor(time),
         bl_level_agreement_function_dummy = as.factor(bl_level_agreement_function_dummy),
         bl_level_agreement_pain_dummy = as.factor(bl_level_agreement_pain_dummy),
         bl_agreement_function = as.factor(case_when(bl_agreement_function == 0 ~ "disagree", 
                                                     bl_agreement_function == 1 ~ "agree")),
         bl_agreement_pain = as.factor(case_when(bl_agreement_pain == 0 ~ "disagree",
                                                 bl_agreement_pain == 1 ~ "agree")))
    
# write_csv(data_mixed_change_tess, "~/Desktop/data_mixed_change_tess.csv")

gamlj::gamljMixed(
    formula = score ~ 1 + time + bl_level_agreement_function_dummy + time:bl_level_agreement_function_dummy+( 1 | subject_id ),
    data = data_mixed_change_tess,
    plotHAxis = time,
    plotSepLines = bl_level_agreement_function_dummy,
    plotError = "ci")
```

### based on agreement on pain

```{r echo=FALSE, fig.width=8}
gamlj::gamljMixed(
    formula = score ~ 1 + time + bl_level_agreement_pain_dummy + time:bl_level_agreement_pain_dummy+( 1 | subject_id ),
    data = data_mixed_change_tess,
    plotHAxis = time,
    plotSepLines = bl_level_agreement_pain_dummy,
    plotError = "ci")
```

## UNIVARIATE: OUTCOME = Quality of Life (VAS scale)

### based on agreement on function

```{r echo=FALSE, fig.width=8}
data_mixed_change_qol <- data %>%
  filter(bl_deceased != "Yes") %>%
  dplyr::select(subject_id, gender, age_y, education_num, marital, tumor_grade, change_function_direction, change_pain_direction, bl_level_agreement_pain, bl_level_agreement_function, bl_agreement_function, bl_agreement_pain, bl_level_agreement_function_dummy, bl_level_agreement_pain_dummy,  
                bl_qol_7, m12_qol_7) %>%
  pivot_longer(cols = c(bl_qol_7, m12_qol_7),
               names_to = "time",
               values_to = "score") %>%
  mutate(subject_id = as.numeric(subject_id),
         time = as.factor(time),
         bl_level_agreement_function_dummy = as.factor(bl_level_agreement_function_dummy),
         bl_level_agreement_pain_dummy = as.factor(bl_level_agreement_pain_dummy),
         bl_agreement_function = as.factor(case_when(bl_agreement_function == 0 ~ "disagree", 
                                                     bl_agreement_function == 1 ~ "agree")),
         bl_agreement_pain = as.factor(case_when(bl_agreement_pain == 0 ~ "disagree",
                                                 bl_agreement_pain == 1 ~ "agree")))
    

# write_csv(data_mixed_change_qol, "~/Desktop/data_mixed_change_qol.csv")

gamlj::gamljMixed(
    formula = score ~ 1 + time + bl_level_agreement_function_dummy + time:bl_level_agreement_function_dummy + ( 1 | subject_id ),
    data = data_mixed_change_qol,
    plotHAxis = time,
    plotSepLines = bl_level_agreement_function_dummy,
    plotError = "ci")
```

### based on agreement on pain

```{r echo=FALSE, fig.width=8}
gamlj::gamljMixed(
    formula = score ~ 1 + time + bl_level_agreement_pain_dummy + time:bl_level_agreement_pain_dummy+( 1 | subject_id ),
    data = data_mixed_change_qol,
    plotHAxis = time,
    plotSepLines = bl_level_agreement_pain_dummy,
    plotError = "ci")
```

## UNIVARIATE: OUTCOME = FATIGUE (FACIT)

### based on agreement on function

```{r echo=FALSE, fig.width=8}
data_mixed_change_facit <- data %>%
  filter(bl_deceased != "Yes") %>%
  dplyr::select(subject_id, gender, age_y, education_num, marital, tumor_grade, change_function_direction, change_pain_direction, bl_level_agreement_pain, bl_level_agreement_function, bl_agreement_function, bl_agreement_pain, bl_level_agreement_function_dummy, bl_level_agreement_pain_dummy, 
                bl_facit_score, m12_facit_score) %>%
  pivot_longer(cols = c(bl_facit_score, m12_facit_score),
               names_to = "time",
               values_to = "score") %>%
  mutate(subject_id = as.numeric(subject_id),
         time = as.factor(time),
         bl_level_agreement_function_dummy = as.factor(bl_level_agreement_function_dummy),
         bl_level_agreement_pain_dummy = as.factor(bl_level_agreement_pain_dummy),
         bl_agreement_function = as.factor(case_when(bl_agreement_function == 0 ~ "disagree", 
                                                     bl_agreement_function == 1 ~ "agree")),
         bl_agreement_pain = as.factor(case_when(bl_agreement_pain == 0 ~ "disagree",
                                                 bl_agreement_pain == 1 ~ "agree")))
    
# write_csv(data_mixed_change_facit, "~/Desktop/data_mixed_change_facit.csv")

gamlj::gamljMixed(
    formula = score ~ 1 + time + bl_level_agreement_function_dummy + time:bl_level_agreement_function_dummy+( 1 | subject_id ),
    data = data_mixed_change_facit,
    plotHAxis = time,
    plotSepLines = bl_level_agreement_function_dummy,
    plotError = "ci")
```

### based on agreement on pain

```{r echo=FALSE, fig.width=8}
gamlj::gamljMixed(
    formula = score ~ 1 + time + bl_level_agreement_pain_dummy + time:bl_level_agreement_pain_dummy+( 1 | subject_id ),
    data = data_mixed_change_facit,
    plotHAxis = time,
    plotSepLines = bl_level_agreement_pain_dummy,
    plotError = "ci")
```


## Post meeting analysis - change over time

Univariate models indicate that the change in function was associated with TESS and fatigue at 12months.

```{r, echo=FALSE, results=FALSE}
data %>%
  group_by(change_function_direction) %>%
  count() %>%
  drop_na()

data %>%
  group_by(change_pain_direction) %>%
  count() %>%
  drop_na()

m_tess_change_function <- lm(m12_tess ~ change_function_direction, data = data)
summary(m_tess_change_function) # YES

m_tess_change_pain <- lm(m12_tess ~ change_pain_direction, data = data)
summary(m_tess_change_pain) # NO
###
m_qol_change_function <- lm(m12_qol_7 ~ change_function_direction, data = data)
summary(m_qol_change_function) # NO

m_qol_change_pain <- lm(m12_qol_7 ~ change_pain_direction, data = data)
summary(m_qol_change_pain) # NO
###
m_facit_change_function <- lm(m12_facit_score ~ change_function_direction, data = data)
summary(m_facit_change_function) # YES

m_facit_change_pain <- lm(m12_facit_score ~ change_pain_direction, data = data)
summary(m_facit_change_pain) # NO


multinomial <- data %>%
  dplyr::select(subject_id,change_pain_direction, change_function_direction, gender, age_y, education_num, marital, tumor_grade, ethnicity)

# write_csv(multinomial, "~/Desktop/multinomial.csv")

data %>%
  dplyr::select(change_function_direction) %>%
  group_by(change_function_direction) %>%
  count() %>%
  drop_na() %>%
  ungroup() %>%
  mutate(sum = sum(n),
         Prop = (prop = round(n/sum,2)*100)) %>%
  dplyr::select(-sum) %>%
  kable() %>%
  kable_styling(full_width = FALSE)

data %>%
  dplyr::select(change_pain_direction) %>%
  group_by(change_pain_direction) %>%
  count() %>%
  drop_na() %>%
  ungroup() %>%
  mutate(sum = sum(n),
         Prop = (prop = round(n/sum,2)*100)) %>%
  dplyr::select(-sum) %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```

